Peer Assessment 1
=================
B. Kelley
March 11-15, 2015

[Data set obtained on March 11, 2015 from this URL](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)


Read in the data from activity.csv after unzipping the download

```{r getdata, echo=TRUE }
        Dataset<- read.table(file="activity.csv",sep=",", header=TRUE, colClasses=c("numeric","character","numeric"))
        
        colnames(Dataset)
        RowsInOriginalDataset<- nrow(Dataset)
```
Now clean Dataset to remove missing values (NA)

```{r cleanDataset, echo=TRUE}
cleanDataset<- Dataset[complete.cases(Dataset),]
cleanDataset$date<- as.Date(cleanDataset$date)
RowsInDatasetNARemoved<- nrow(cleanDataset)
```
Calculate the total number of steps taken per day
and plot a histogram

```{r sumByday, echo=TRUE}
library("dplyr")
sumByDay<- group_by(cleanDataset,date) %>% summarize(Nsteps=sum(steps))
hist(sumByDay$Nsteps,breaks=length(unique(cleanDataset$date)),
     main="Total Number of Steps Each Day", xlab="Steps",ylab="Number of Days",
     col=rainbow(10))
Meansteps<- mean(sumByDay$Nsteps)
Mediansteps<- median(sumByDay$Nsteps)
```

Mean steps per day is `r Meansteps`. Median steps per day is `r Mediansteps`.

Now get average of steps across all days for each interval

```{r intervalMean, echo=TRUE}
intervalMeans<- group_by(cleanDataset, interval) %>% 
        summarize(avSteps= mean(steps)) %>% arrange(interval)

plot(intervalMeans$interval, intervalMeans$avSteps, xlab="Interval (5 min increments)", ylab="Average Steps",type="l", xlim= c(0,2400))
```
```{r missingvaluesrows, echo=TRUE}
   numRowswithNA<- RowsInOriginalDataset - RowsInDatasetNARemoved
```


The number of rows with missing values is: `r numRowswithNA`

Use the mean steps per day per interval (Meansteps/288) to estimate missing values
 
```{r fillNA, echo=TRUE}
        NArows<- which(is.na(Dataset))
        filleddataset<- Dataset
        filleddataset$steps[NArows]<- Meansteps/288

        sumByDay<- group_by(filleddataset,date) %>% summarize(Nsteps=sum(steps))
        hist(sumByDay$Nsteps,breaks=length(unique(filleddataset$date)),
                main="Total Number of Steps Each Day", xlab="Steps",ylab="Number of Days",
        col=rainbow(10))
        Meanstepsfill<- mean(sumByDay$Nsteps)
        Medianstepsfill<- median(sumByDay$Nsteps)
```
With NA's filled in with Meansteps/total number of intervals:
Mean steps per day is `r Meanstepsfill`. Median steps per day is `r Medianstepsfill`.

With this simple strategy, the mean did not change and the median shifted to equal the mean.

Now plot by part of week (weekday/weekend)

```{r dayofweekfactors , echo=TRUE}
        Weekday<- c("Monday","Tuesday","Wednesday","Thursday","Friday")
        Weekend<- c("Saturday","Sunday")
        filleddataset<- mutate(filleddataset, partOfweek= weekdays(as.Date(filleddataset$date)))
        weekday<- filter(filleddataset, partOfweek == "Monday"|partOfweek=="Tuesday"|partOfweek=="Wednesday"
                         |partOfweek=="Thursday"|partOfweek=="Friday")
        weekend<- filter(filleddataset, partOfweek == "Saturday"|partOfweek== "Sunday")
        
        weekdayMeans<- group_by(weekday, interval) %>% 
                summarize(avSteps= mean(steps)) %>% arrange(interval)
        weekendMeans<- group_by(weekend, interval) %>%
                summarize(avSteps= mean(steps)) %>% arrange(interval)

        par(mfcol= c(2,1) )
        plot(weekdayMeans$interval, weekdayMeans$avSteps, ylab="Average Steps",xlab="Interval", 
             type="l", xlim= c(0,2400), main="weekday")
        plot(weekendMeans$interval, weekendMeans$avSteps, ylab="Average Steps",xlab="Interval",
             type="l", xlim= c(0,2400), main="weekend")
```

